### 2023-02-16 doc/manuscript/translocation.qmd
Minor edits to comments

Host evolution following disease-induced declines is one mechanism by which hosts can recover [@carlson2014; @searle2020]. Evolutionary rescue describes natural host evolution whereby rapid evolutionary changes originating from variation within a host population reverse ongoing population declines [@carlson2014]. In host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes [@golas2021]. Evolutionary changes following disease-induced declines have been repeatedly observed in wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], but linking these selective responses directly to population recoveries, as necessary for evolutionary rescue, remains rare. Important for conservation action, selection that promotes recoveries of focal populations can also provide sources of resistant individuals to reestablish populations extirpated following disease emergence [@joseph2018; @mendelson2019]. These so-called "recovery engines" [@mendelson2019] -- where rescued (i.e., resistant and recovered) host populations are used to successfully repatriate extirpated populations on the landscape -- are an intriguing management idea, but have not been successfully implemented in wildlife conservation efforts.

@@

![Map showing the locations of translocated and donor populations in Yosemite National Park (park boundary indicated by black polygon). Symbol shapes indicate the donor population used for each translocation site. To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of Yosemite within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation). ](figures/map_translocation_points.png){#fig-yosemap width="450"}

>>>>

Host evolution following disease-induced declines is one mechanism by which hosts can recover [@carlson2014; @searle2020]. Evolutionary rescue describes natural host evolution whereby rapid evolutionary changes originating from variation within a host population reverse ongoing population declines [@carlson2014]. In host-pathogen systems, evolutionary rescue is more likely to occur when hosts experience repeated selection events (e.g., through re-infections), there is a persistent pathogen reservoir or frequency-dependent infection dynamics, and populations experience increased per capita birth rates at small population sizes [@golas2021]. Evolutionary changes following disease-induced declines have been repeatedly observed in wildlife populations [@savage2016; @epstein2016; @gignoux-wolfsohn2021; @holland2022], but linking these selective responses directly to population recoveries, as necessary for evolutionary rescue, remains rare. Important for conservation action, selection that promotes recoveries of focal populations can also provide sources of resistant individuals to reestablish populations extirpated following disease emergence [@joseph2018; @mendelson2019]. These so-called "recovery engines" [@mendelson2019] -- where rescued (i.e., resistant and recovered) host populations are used to successfully repatriate extirpated populations on the landscape -- are an intriguing management idea, but have not been successfully implemented in wildlife conservation efforts. <!--# Roland says: Modestly speaking, Joseph and Knapp (2018) does provide an example, and one that predates Mendelson (2019) paper. -->

@@

![Map showing the locations of translocated and donor populations in Yosemite National Park (park boundary indicated by black polygon). Symbol shapes indicate the donor population used for each translocation site. To obscure the exact locations of populations, random noise was added to all point coordinates. Inset map shows the location of Yosemite within California. In both maps, elevation is indicated by the colored hillshade layer (dark green = lowest elevation, white = highest elevation).](figures/map_translocation_points.png){#fig-yosemap width="450"}
